---
title: "p8105_hw2_xs2569"
author: "Xun Sun"
date: "2024-09-26"
output:
  github_document: null
  html_document:
    df_print: paged
---

# 1] problem 1  -  NYC Transit data

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

### 1.1 read and clean
```{r import}
subway_df = 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na = c("NA", "", ".")) |>
  janitor::clean_names() |> 
  mutate( 
  entry_logic = ifelse(is.na(entry), FALSE, TRUE),
  route_count = 11 - rowSums(is.na(across(starts_with("route"), .names = "route")))) |> 
  select(line, station_name, station_latitude, station_longitude, route_count, entry_logic, vending, entrance_type, ada)

view(subway_df)
summary(subway_df)
```

##### This data frame :subway_df) now contains 9 variables as follows:
- **line** (`Character`): Indicates the subway line. 
- **station_name** (`Character`): The name of the subway station. 
- **station_latitude** (`Numeric`): The latitude coordinate of the station. Values range from 40.58 to 40.90, with a median at 40.73.
- **station_longitude** (`Numeric`): The longitude coordinate of the station. Values range from -74.03 to -73.76, with a median at -73.96.
- **route_count** (`Numeric`): The number of routes available at each station (calculated). On average, stations have about 2.29 routes available, ranging from 1 to 11.
- **entry_logic** (`Logical`): A logical indicator of whether there is an entry (`TRUE` if there is an entry, `FALSE` otherwise). 468 stations have entries marked as `TRUE`, while 1400 are marked as `FALSE`.
- **vending** (`Character`): Information about vending availability at the station. The variable contains descriptive data about vending services.
- **entrance_type** (`Character`): Type of entrance. This variable describes the type of entrance each station has.
- **ada** (`Logical`): Logical indicator for whether the station is ADA accessible (`TRUE` if accessible, `FALSE` otherwise). 468 stations are ADA accessible, while 1400 are not.

Describe data cleaning steps:
1. **Read Data**: Import the dataset, treating specific strings as NA values.
2. **Clean Column Names**: Standardize column names for consistency.
3. **Calculate Entry Logic**: Create a binary indicator for entry presence.
4. **Determine Route Count**: Compute the number of non-missing route entries per station.
5. **Select Columns**: Keep only the relevant columns for further analysis.

The dimension of the resulting dataset:
The resulting data set contains 9 columns and 1868 rows. 

After the steps above, the data can still be cleaner.

### 1.2 answer questions
How many distinct stations are there?
```{r}

```

#Problem 2
